\documentclass{article}
\usepackage{listings}
\usepackage[colorlinks=true,linkcolor=black,anchorcolor=black,citecolor=black,filecolor=black,menucolor=black,runcolor=black,urlcolor=black]{hyperref}
\usepackage{pythonhighlight}
\usepackage{graphicx}
\setlength{\parindent}{0pt}

\begin{document}
	\title{Python: Data Visualization Notes}
	\author{Klein \\ carlj.klein@gmail.com}
	\date{}
	\maketitle
	
\section{Learning Objectives}
What good is a data analysis if the answers or findings can't be conveyed properly? Or perhaps even more detrimental, what good is a dataset if the proper questions or hypotheses can't be formed in the first place? How can we properly tell the story the data is providing? Enter in data visualization! Both explaining the data and exploring the data can be significantly helped through the process of data visualization.
\\\\
Data visualization is two-pronged:
\begin{itemize}
	\item Exploratory Analysis: Helping to understand the data prior to analysis. Searching for relationships and insights.
	\item Explanatory Analysis: Helping to present analysis findings, helping to tell a story with the data. After insights were found.
\end{itemize}


And it's all a part of the entire data analysis process. We can also simplify the data analysis process into 5 steps:
\begin{enumerate}
	\item Extract
	\item Clean: Exploratory
	\item Explore: Exploratory
	\item Analyze: Exploratory OR Explanatory
	\item Share: Explanatory
\end{enumerate}

In this course, we'll be using the matplotlib, seaborn, and Pandas libraries to assist in the data analysis process.
\\\\

Concepts
\begin{itemize}
	\item \nameref{sec:concept1}
	\item \nameref{sec:concept2}
		\begin{itemize}
			\item \nameref{sec:concept2.1}
			\item \nameref{sec:concept2.2}
			\item \nameref{sec:concept2.3}
		\end{itemize}
	\item \nameref{sec:concept3}
	\item \nameref{sec:concept4}
\end{itemize}
	
\section{Design of Visualization}\label{sec:concept1}
\input{Concept1}

\section{Exploration of Data}\label{sec:concept2}
We have a dataset on a topic or concept that has been deemed worthy for inspection! Surely, there are some insights to be gained from it. We load up the data, and then we hit a wall... Which columns are important? What questions can be answered? We can find the general statistics of the set, so what?
\\\\

This section will help with the initial process of data exploration, helping to find what is actually useful and should be examined further in the data. We'll start with single variables from the data and move into visually pairing multiple data points at once.
\\\\

We'll be using a few different Python libraries in this section:

\begin{python}
	import numpy as np
	import pandas as pd
	import matplotlib.pyplot as plt
	import seaborn as sb
\end{python}

\subsection{Univariate Exploration of Data}\label{sec:concept2.1}
\input{Concept2.1}

\newpage
\subsection{Bivariate Exploration of Data}\label{sec:concept2.2}
\input{Concept2.2}

\newpage
\subsection{Multivariate Exploration of Data}\label{sec:concept2.3}
\input{Concept2.3}

\newpage
\section{Explanatory Visualizations}\label{sec:concept3}
\input{Concept3}

\section{Visualization Case Study}\label{sec:concept4}
\input{Concept4}
	
\end{document}